Night and day: What is the influence and relative importance of urban characteristics on land surface temperature?

### 2019-05-21 report/article.tex
updating with katie's edits and andrew's appendix

% \documentclass[sort&compress]{elsarticle}
\documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
%\documentclass[review]{elsarticle}

@@

%% Packages
\usepackage{tabu}
\usepackage{breakurl}
\usepackage{float}

@@

\title{Night and day: What are the nonlinear associations between urban characteristics and land surface temperature?}

@@

\begin{abstract}
Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
Understanding these drivers is necessary for reducing land surface temperature and mitigating the effects of heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
% However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
However, the independent effects and relative importance of potential drivers remain unclear from existing studies.
We seek to answer the question: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?'' 
To do this, we analyze urban land surface temperature in four cities across the United States.
In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
We model the effects of these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
The effects from the daytime and nighttime analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
Our results suggest that vegetation and impervious surfaces are the most important urban characteristics associated with land surface temperature.
Increasing and decreasing these, respectively, is necessary for reducing high urban temperatures during both night and day.
Our results also demonstrate the potential for using nonlinear statistical analysis to investigate land surface temperature and its relationships with urban characteristics. 
Improved understanding of these relationships influencing both night and day land surface temperature will assist planners undertaking climate change adaptation and heat wave mitigation.
\end{abstract}

@@

\section{Introduction}
In a warming world, understanding the drivers of high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and well being of their communities.
And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}. 
Heat waves' affect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
The UHI results from heat captured during the day being released during the night, thus increasing nocturnal temperatures \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
This increase in nocturnal temperature reduces people's ability to cool off during the night and drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.

Urban heat is studied in two ways.
The first uses weather station data to analyze air temperature (e.g. \cite{Scott2016-lc}) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
Although there are differences between LST and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
We analyze LST in this study.

@@

Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}. 
Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relax their linear assumption, 4) rigorously testing and validing their models.
The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
1km resolution makes it difficult to attribute LST to urban characteristics.
However, 30-m resolution LandSat8 (L8) night scenes have recently become available. 
Due to the wide spatial availability of L8 imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.

Existing studies have also been criticized for the explanatory variables they've used \cite{Chun2017-mm,Peng2018-cp}.
There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
Beyond the 2D or 3D debate, important variables include green space, water bodies, albedo, and socio-economic factors \cite{Peng2018-cp}. 
However, many studies do not capture these categories and, worse, many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion). 
Considering a variable in isolation, without accounting for potential conflating by other variables, limits the understanding of the interdependent effects that exist.

The third potential enhancement is in the statistical models.
Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
Using linear models incur a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
The first is that many of the urban characteristics exhibit high multicollinearity \cite{Zhou2014-wc}.
The second is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \cite{Peng2018-cp, Zhou2014-wc}.

Finally, and perhaps most importantly, the existing studies fail to appropriately validate their models.
The approach they use to assess their accuracy is in-sample validation.
In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data. 
This risks overestimating the accuracy of their models.

The question we address is: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?''
To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
In addressing our question, we will address the, at times fundamental, limitations of existing studies.
Most importantly we look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
The results have implications for mitigating the severity of future heat waves.

@@

Figure \ref{fig:map} shows the gridded LST data for each of the cities and Figure \ref{fig:joy} shows the distribution of each city's day and night land surface temperature.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/report/map_nocturnal_lst.pdf}
    \caption{The gridded nocturnal land surface temperature in $^o$C of the cities studied. This is change from the mean (average) for each city.}
    \label{fig:map}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fig/report/joyplot_lst_500.pdf}
    \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied.}
    \label{fig:joy}
\end{figure}

\subsection{Land surface temperature}
We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}). 
Our code is available on our Github\footnote{URL redacted for review} repository, and is as follows:
We convert Band-10 digital-number data to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. 
We correct for emissivity using land cover data \cite{Alipour2003-gb}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
This follows the process is described in \cite{Scott2016-lc}.

To ameliorate the effect of shadows and other ephemeral changes \cite{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{fig/report/scatter_500.png}
    \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution.}
    \label{fig:scatter_lst}
    \end{center}
\end{figure}

@@

\textit{Impervious surface percentage}. This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}. To generate the impervious surface area (ISA) LandSat data, NLCD land cover, and nighttime light imagery is used. The stable nighttime light intensity is only used to estimate the boundary of of urban areas. The Landsat images were converted to top-of-atmosphere reflectance. The data is provided at a 30m resolution.

@@

\textit{NBDI}. 
Normalized difference built-up index indicates the intensity of imperviousness \cite{Bhatti2014-ae}. 
It is calculated from satellite images as $$NDVI=\frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}}$$ where $B_{NIR}$ and $B_{SWIR}$ are the reflectances in the near-infrared and short-wave infrared bands respectively \cite{Alhawiti2016-wv}. 
Using LandSat8 imagery, this is $$NDVI=\frac{B6-B5}{B6+B5}$$ \cite{barsi2014}.
% *Kaplan 2018, Peng 2018

@@

\textit{Population density}. 
Density allegedly increases the LST \cite{Li2017-yl, Peng2018-cp}, although this result may be the result of confounding with other factors. 
Population density is calculated from the U.S. census at the block level. 
The most recent census was 2010, so we use that data as an estimator of where people reside in the evening. 
This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that it's centroid is contained within.

\textit{Sky view factor}. Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}. We calculate SVF using \cite{Van_doninck2018-ib} and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. We use a spatial resolution of 6 meters.
% * Yuan 2011, Unger 2004.

\textit{Tree canopy cover}. 
The percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}. 
The data provided is at a 30m resolution. 
The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance). 
% * Rogan 2013, Elmes 2017

\subsection{Data preparation and robustness}
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
    \caption[Holdout cross-validation results at 500 meter resolution]{
    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
    The models were trained on 80\% of the data and tested on the unseen 20\%.
    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
    Good models have MAE near 0.
    }
    \label{fig:holdout_500}
\end{figure*}

The objective of our study is to determine the most influential urban characteristics and understand how they relate to land surface temperature.
To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
Given a dataset and for a chosen characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature. 
This is repeated over a range for the feature of interest.
To evaluate how multiple urban characteristics interact, partial dependence can also be conducted in multiple dimensions.
The swing \cite{Shortridge2015-ub} measures the relative importance of variables by calculating the magnitude of change in temperature associated with each variable.
To calculate partial dependence, we need a data set and a statistical model.

To analyze the data, it needs to be transformed into a suitable format that is consistent between the variables.
The raw data is a variety of spatial data types from ``area level" (the population density data), to ``geostatistical raster" (e.g. the land surface temperature).
We choose to resample the data into a grid of square cells.
We conduct the resampling twice to produce a grid of 500m and 100 meter cells. 
Having two grid resolutions allowed us to ensure our conclusions are robust.
For each of the cells the the mean, maximum, and minimum of all variables were calculated. 
To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable. 
To address multicollinearity between the variables, that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than approximately 5.
The result is a data set which with we can train a variety of statistical models.

Various statistical models (\S \ref{ss:models}) were trained and then, crucially, validated on unseen data using a technique known as holdout cross-validation. 
This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
The model resulting from the training, is then tested on this unseen test set (known as out-of-bag), to get an estimate of the model accuracy.
This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
Over fitting occurs when a model fits to the randomness in a dataset, causing it to be unsuitable for generalizations. 
When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
Avoiding over fitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies (e.g. \cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
Cross-validation helps to avoid this by evaluating the model on unseen data.

A further way to avoid over fitting is to carefully consider how the test data is selected.
This is especially important with spatial data.
Therefore, when selecting the test and training data the grid cells were grouped into a larger 8x8 grid.
This avoids over fitting of the model by training the model on a cell adjacent to a cell that is included in the test set.

Finally, we attempt to capture uncertainty in the data and the models. 
To represent the data uncertainty we draw a random sample from our data, with replacement, to simulate a new dataset.
The models are trained on this new dataset and this is repeated.
This approach, known as bootstrapping, means that the conclusions' sensitivity to the data can be assessed.
If the results are vastly different for each bootstrap sample, we assume that the result is dependent on the data and is unlikely to be generalizable to other cities.
Additionally, we capture model uncertainty by using different models and assessing how each model shows the urban characteristics influencing the land surface temperature.
If all bootstrapped models are generally in agreement, it suggests that the conclusions are robust.

@@

\textit{Linear model}. Linear models are suitable when there is a linear relationship between the explanatory variables (urban characteristics) and the response variable (LST).

\textit{Multivariate Adaptive Regression Spline (MARS)}. The MARS model extends the linear model, by using piece-wise functions to fit the data. \cite{Friedman1991-of}

\textit{Generalized Additive Model (GAM)}. The GAM is also an extension of the linear model, although it does not require a linear relationship. Instead, the response variable is estimated as the sum of smoothing functions that are applied to each covariate or set of covariates \cite{Hastie1990-cg}.

\textit{Random Forest}. A random forest is an ensemble model of \textit{Regression Trees}. 
A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}. It continues to divide the data on different covariates until the partitions are as similar as possible. 
The result is a tree-like structure, after which the model is named.
In similar naming fashion, a random forest is a collection of regression trees.
Each tree is \textit{grown} from a random subset of the data and the prediction from the random forest model is the average of the prediction from all of the trees \cite{Breiman2001-rt}.
Tree-based models do not assume linearity and so are generally very flexible and powerful models \cite{Breiman2001-rt, Geron2017-ek}.

\textit{Gradient Boosted Regression Trees}. This is similar to the random forest, that it is a collective of regression trees. 
However, the difference is that each tree is trained sequentially on the residuals of the previous tree.
The result is the average of all of the regression trees \cite{Geron2017-ek}.

\textit{Convolutional Neural Network}. 
<add explanation>

\section{Results and Discussion}
In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
To achieve this, the statistical models need to be accurate.
The results in Figure \ref{fig:holdout_500} indicate that both day and night land surface temperature can be predicted within 1$^o$C at night and 2$^o$C during the day using urban characteristics.
The R$^2$ results, also based on unseen data, suggest that more than 80\% of the data variance is captured by the models for the mean LST during both night and day, while the mean \% variance explained is generally between 50 and 80\% for the maximum temperatures.
Compared to models presented in similar studies, these are strong results, especially given that the existing studies failed to validate their models on unseen data.
We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees.
The weakest model is the linear regression, incidentally the one that is used in the majority of existing studies into land surface temperature.
This strong result allows us to assess the influence and relative importance that these characteristics.

\begin{figure*}
    \begin{center}
    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
    \caption[Variable influence on LST at 500 meter resolution]{
    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
    \label{fig:importance_500}
    \end{center}
\end{figure*}

>>>>

% \documentclass[sort&compress]{elsarticle}
% \documentclass[final,3p,times,onecolumn,sort&compress]{elsarticle}
\documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
%\documentclass[review]{elsarticle}

@@

%% Packages
\usepackage{tabu}
\usepackage{breakurl}
\usepackage{float}
\usepackage{amsmath}

@@

\title{Night and day: What is the influence and relative importance of urban characteristics on land surface temperature?}

@@

\begin{abstract}
Heat waves are among the deadliest natural hazards and are expected to increase in frequency and severity under climate change. 
Their impacts in cities can be exacerbated by the urban heat island (UHI), but the mechanisms that underlie severity and timing of the UHI and its interactions with heat waves are poorly understood.
Understanding these mechanisms is necessary to design strategies that reduce land surface temperature and mitigate the effect of heat waves.
Making use of recently available high-resolution day and night thermal satellite imagery and employing advanced nonlinear statistical models, we seek to answer the question \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''} 
To answer this question, we analyze urban land surface temperature in four cities across the United States.
In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
We model the effects of these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
The effects from the daytime and nighttime analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
Our results suggest that vegetation and impervious surfaces are the most important urban characteristics associated with land surface temperature.
Increasing and decreasing these, respectively, is necessary for reducing high urban temperatures during both night and day.
Our results also demonstrate the potential for using nonlinear statistical analysis to investigate land surface temperature and its relationships with urban characteristics. 
Improved understanding of these relationships influencing both night and day land surface temperature will assist planners undertaking climate change adaptation and heat wave mitigation.
\end{abstract}

@@

\section{Introduction}
In a warming world, understanding the factors contributing to high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and wellbeing of their communities.
And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}. 
Heat waves' effect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
The UHI is understood to be a product of multiple factors including: enhanced absorption of solar radiation, geometric effects that limit radiative cooling and ventilation, anthropogenic heat from vehicles and buildings, air pollution that traps outgoing radiation, high ratios between sensible and latent heat flux due to low vegetation cover and high impervious land cover, and the radiative characteristics of building materials. % \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
The result is increased diurnal and nocturnal temperatures that reduce people's ability to cool off, especially during the night, which drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.

Studies of urban heat include analyses of air temperature, which is a conventional meteorological variable that is often associated with health outcomes(e.g. \cite{Scott2016-lc}), and studies that use satellite-derived land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
LST correlates with air temperature at large scales, but there are differences at intra-urban scales \cite{Good2016-yk}.
While LST is not a conventional meteorological variable, it is directly related to the urban heat budget and has the advantage of being available with extensive spatial coverage in gridded satellite products \cite{Hung2006-qy}.
We analyze LST in this study.

@@

Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}. 
Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relaxing their linear assumption, 4) rigorously testing and validating their models.
The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
1km resolution makes it difficult to attribute LST to urban characteristics.
However, 100m resolution LandSat8 (L8) night scenes have recently become available. 
Due to the wide spatial availability of L8 imagery, comparative studies between cities that are necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.

Existing studies have also been criticized for the explanatory variables they've used \cite{Chun2017-mm,Peng2018-cp}.
There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
Beyond the 2D or 3D debate, important variables include green space, water bodies, albedo, and socio-economic factors \cite{Peng2018-cp}. 
However, many studies do not capture these categories and many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion). 
Considering a variable in isolation, without accounting for potential conflating by other variables, limits the understanding of the interdependent effects that exist.

The third potential enhancement is in the statistical models.
Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
Using linear models incurs a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
The first is that many of the urban characteristics exhibit high multicollinearity \cite{Zhou2014-wc}.
The second is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \cite{Peng2018-cp, Zhou2014-wc}.

Finally, most existing studies do not rigorously validate their models.
The approach they use to assess their accuracy is in-sample validation.
In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data. 
This risks overestimating the accuracy of their models.

The question we address is: \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
In addressing our question, we will address the, at times fundamental, limitations of existing studies.
We look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
The results have implications for mitigating the severity of future heat waves.

@@

Figure \ref{fig:map} shows the 500m gridded nighttime LST data for each of the cities and Figure \ref{fig:joy} shows the distribution of each city's 100m resolution day and night land surface temperature.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/report/map_nocturnal_lst.pdf}
    \caption[The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied]{The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied. This is change from the mean (average) for each city.}
    \label{fig:map}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fig/report/joyplot_lst_100.pdf}
    \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied. 100m resolution.}
    \label{fig:joy}
\end{figure}

\subsection{Land surface temperature}
We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}). 
Our code is available on our Github\footnote{URL redacted for review} repository, and is as follows:
We convert Band-10 digital-number data to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. 
We correct for emissivity using land cover data \cite{Alipour2003-ym}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
Finally, an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
This follows the process is described in \cite{Scott2016-lc}.

To ameliorate the effect of ephemeral changes \cite{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{fig/report/scatter_100.png}
    \caption{The land surface temperature in $^o$C of the cities studied at a 100m square resolution.}
    \label{fig:scatter_lst}
    \end{center}
\end{figure}

@@

\textit{Impervious surface percentage}. This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}. 
To generate the impervious surface area (ISA), the consortium uses LandSat data, NLCD land cover, and nighttime light imagery. 
The stable nighttime light intensity is only used to estimate the boundary of urban areas. 
The Landsat images were converted to top-of-atmosphere reflectance. 
The data is provided at a 30m resolution.

@@

\textit{NDBI}. 
Normalized difference built-up index indicates the intensity of imperviousness \cite{Bhatti2014-ae}. 
It is calculated from satellite images as $$NDBI=\frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}}$$ where $B_{NIR}$ and $B_{SWIR}$ are the reflectances in the near-infrared and short-wave infrared bands respectively \cite{Alhawiti2016-wv}. 
Using LandSat8 imagery, this is $$NDBI=\frac{B6-B5}{B6+B5}$$ \cite{barsi2014}.
% *Kaplan 2018, Peng 2018

@@

\textit{Population density}. 
Some studies have found that population density has a positive influence on LST \cite{Li2017-yl, Peng2018-cp}, although this may be the result of confounding with other factors. 
We calculate population density using the U.S. census at the block level. 
The most recent census was 2010, so we use that data as an estimator of where people reside in the evening. 
This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that its centroid is contained within.

\textit{Sky view factor}. Urban canyons have been found to have an effect on UHI because they prevent air circulation and hinder nighttime radiative cooling \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}. We calculate SVF using the R packages \textit{horizon} \cite{Van_doninck2018-ib} that is based off the algorithm presented in \cite{Dozier1990-kn}. 
We use the same parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. 
We use a spatial resolution of 6 meters.
% * Yuan 2011, Unger 2004.

\textit{Tree canopy cover}. 
The 2011 edition of percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}. 
The data provided is at a 30m resolution. 
The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance). 
% * Rogan 2013, Elmes 2017

\subsection{Data preparation and robustness}
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{fig/report/holdout_100.png}
    \caption[Holdout cross-validation results at 100-meter resolution]{
    Holdout cross-validation results at 100-meter resolution.
    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
    The models were trained on 80\% of the data and tested on the unseen 20\%.
    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
    Good models have MAE near 0.
    }
    \label{fig:holdout_100}
\end{figure*}

The objective of our study is to determine the most influential urban characteristics and understand how they relate to land surface temperature.
To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
For a selected characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature. 
This is repeated over a range for the feature of interest.
To evaluate how multiple urban characteristics interact, partial dependence can also be conducted in multiple dimensions.
The swing \cite{Shortridge2015-ub} measures the relative importance of variables by calculating the magnitude of change in temperature associated with each variable.
To calculate partial dependence, we need a data set and a statistical model.

To analyze the data, it needs to be transformed into a suitable format that is consistent between the variables.
The raw data is a variety of spatial data types from ``area level" (the population density data), to ``geostatistical raster" (e.g. the land surface temperature).
We choose to resample the data into a grid of square cells.
We conduct the resampling twice to produce a grid of 100 and 500-meter square cells.

Having two resolutions allowed us to evaluate whether our conclusions are robust.
For each of the cells the mean, maximum, and minimum of all variables were calculated. 
To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable. 
To address multicollinearity between the variables that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than 5.
The result is a data set with which we can train statistical models.

Various statistical models (\S \ref{ss:models}) were trained and then validated on unseen data using a technique known as holdout cross-validation. 
This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
The model resulting from the training is then tested on this unseen test set (known as out-of-bag) to get an estimate of the model accuracy.
This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
Over fitting occurs when a model fits to the randomness in a dataset, causing it to be unsuitable for generalizations. 
When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
Avoiding overfitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies.
Cross-validation helps to avoid this by evaluating the model on unseen data.

A further way to avoid overfitting is to carefully consider how the test data is selected.
This is especially important with spatial data.
Therefore, when selecting the test and training data the grid cells were grouped into a larger 8x8 grid.
This avoids overfitting of the model by training the model on a cell adjacent to a cell that is included in the test set.

Finally, we attempt to capture uncertainty in the data and the models. 
To represent the data uncertainty, we draw a random sample from our data, with replacement, to simulate a new dataset.
The models are trained on this new dataset and this is repeated.
This approach, known as bootstrapping, means that the conclusions' sensitivity to the data can be assessed.
If the results are vastly different for each bootstrap sample, we assume that the result is dependent on the data and is unlikely to be generalizable to other cities.
Additionally, we capture model uncertainty by using different models and assessing how each model shows the urban characteristics influencing the land surface temperature.
If all bootstrapped models are generally in agreement, it suggests that the conclusions are robust.

@@

\textit{Linear model}. Linear models are suitable when the relationship between the explanatory variables (urban characteristics) and the response variable (LST) is linear. 
That is, linear models take the form $$ y = \beta_0 + \sum_i \beta_i x_i + \varepsilon$$ where $y$ is the response (LST), $x_i$ is each for the variables (urban characteristics), $\beta_0$ is the intersection, $\beta_i$ is how the LST changes linearly with each urban characteristic, and $\varepsilon$ is the normally distributed error.

\textit{Multivariate Adaptive Regression Spline (MARS)}. 
MARS models are a type of non-parametric regression approach that are useful for high dimensional problems \cite{Hastie2009-ky}. 
They extend the linear model by using piecewise functions to fit the data \cite{Friedman1991-of}. 
Multiple linear approximations are summed across the variable space to construct an estimate of the  LST\cite{Friedman1991-of}.

\textit{Generalized Additive Model (GAM)}. 
The GAM is also an extension of the linear model but relaxes the assumption that the relationship is linear \cite{Hastie1990-cg}.
Instead, the response variable is estimated as the sum of smoothing functions, splines, that are applied to each covariate or set of covariates: $$ y = \sum_i s_i(x_i)$$ where $s_i(x_i)$ is the smoothing function applied to each covariate \cite{Shortridge2015-ub}.

\textit{Random Forest}. A random forest is an ensemble model of \textit{Regression Trees}. 
A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}. 
The tree continues to `grow' by recursively partitioning the data with the objective of maximizing the node impurity, so that the partitions are as similar as possible. 
The result is a tree-like structure that uses thresholds on the covariates to estimate the response.

We use a total of 500 regression trees.
Each regression tree is trained on a randomly selected subset of 1/3 of the covariates. 
The resulting prediction from the random forest is the unweighted mean of the prediction from all of the trees \cite{Breiman2001-rt}.
A subset of the covariates is used to reduce the correlation between the predictions of each tree.
Tree-based models do not assume linearity and so are generally very flexible and powerful models \cite{Breiman2001-rt, Geron2017-ek}.

\textit{Gradient Boosted Regression Trees}. 
GBRT's are also a collective of regression trees. 
In contrast to random forest models, each tree is trained sequentially on the residuals of the previous tree.
That is, each regression tree has the objective of reducing the error of the previous tree.
The prediction from the GBRT is the unweighted mean of all of the regression trees \cite{Geron2017-ek}.
In this case, we again use 500 trees.
